---
title: "Exploring Graphics"
output:
  html_document: default
  html_notebook: default
---

# Base R

```{r}
mlb <- read.csv2("../../data/mlb.csv")
head(mlb)
```

## Scatter
```{r}
hist(mlb$height)
```


## histogram

```{r}
hist(mlb$Age)
```




## ggplot2
```{r}
library(ggplot2)
g <- ggplot(data=mlb)
g + geom_bar(aes(x = height)) +
  theme_bw() + 
  labs(title = "Player size")
```

```{r}
g + geom_point(aes(x = height,
                   y = weight,
                   col = Position)
               ) +
  labs(title = "Size vs. weight")
```

## Interactive (plotly)

```{r}
library(plotly)
p <- plot_ly(mlb,
  x = ~height,
  y = ~weight,
  name = "SF Zoo",
  type = "scatter"
)
p

```

## Interactive (dygraphs)
<!-- https://rstudio.github.io/dygraphs/gallery-custom-plotters.html -->

```{r}
baro_ch <- read.csv2("https://datenservice.kof.ethz.ch/api/ts?keys=kofbarometer&mime=csv",
                     header=T,sep=",",
                     stringsAsFactors = F,
                     dec=".")
class(baro_ch)
head(baro_ch)
str(baro_ch)
baro_ts <- ts(baro_ch$kofbarometer,
              start(1991,1),
              frequency = 12)
library(dygraphs)
dygraph(baro_ts, main = "KOF Barometer") %>% 
  dyRangeSelector()
```
**Hacking dygraphs (Advanced Example)**
```{r}
library(dygraphs)
dyBarChart <- function(dygraph) {
  dyPlotter(dygraph = dygraph,
            name = "BarChart",
            path = system.file("examples/plotters/barchart.js",
                               package = "dygraphs"))
}

dygraph(baro_ts, main = "KOF Barometer") %>% 
  dyRangeSelector() %>%
  dyBarChart()


```





## more examples
We use the big5 dataset from personality-testing.info again.
We limit the dataset to a few countries only for illustration purposes. 

```{r}
big5 <- read.csv2("../../data/big5.csv",sep="\t")

# let's select some countries... 
# mention the :: syntax
big5$country <- plyr::revalue(big5$country,
              c("US" = "USA",
                "DE" = "Germany",
                "CA" = "Canada",
                "CN" = "China",
                "ES" = "Spain",
                "FR" = "France")
)

big5_sub <- big5[big5$country %in% c("USA","Germany","Canada"),]
big5_sub <- big5_sub[big5_sub$gender %in% c(1,2),]


```


Let's explore some of the vizualiation that Hadley Wickham's ggplot2 provides. Note that we create a ggplot object first and 
use varying geoms with the same object later. 


```{r}
require(ggplot2)
countries <- ggplot(data = big5_sub) 

# the point geom
countries + geom_point(aes(country,age))

# maybe we should remove ugly outliers...
big5_sub <- subset(big5_sub,age < 120)
countries <- ggplot(data = big5_sub) 

# the point geom
countries + geom_point(aes(country,age))
```

Again we can use the **countries** object and 
combine it with other geoms.

```{r}
# the layer by layer thing... 
c_h <- countries +
  geom_histogram(aes(x = age,
                     y = ..density..),
                 binwidth=1,
                 colour="black",
                 fill="white")
c_h
# add a density
c_h_d  <- c_h +
  geom_density(aes(x = age),
               alpha = .2,fill = "green")
c_h_d

c_h_d_m <- c_h_d + 
  geom_vline(aes(xintercept = mean(age, na.rm=T)),
             color="red", linetype="dashed", size=1)
c_h_d_m
```


# Task 1
Create a density plot withouth the histogram but with the x-intercept. Use a dotted blue line for the intercept! 


# The Bar Geom

```{r}
country_bar <- countries +
  geom_bar(aes(x = as.factor(E1),fill = factor(race))) +
  # scales free just means that scales can very across
  # facets, default is fixed
  facet_wrap(~ country,scales = "free") + 
  scale_fill_grey()

country_bar
```



# Boxplot geom
Note the importance of factor() declaration here. 

```{r}
country_box <- countries + 
  geom_boxplot(aes(x = factor(country),
                             y = age,
                   fill = factor(gender)))

country_box

```

# Task2
Change country from x aesthetic to facet and plot the same boxplot.


# Aggregation of Data using dplyr
**dplyr** is another R package by Hadley Wickham. It's the high performance follow-up of **plyr**. Also **dplyr** provides a more intuitive interface and reminds of query languages. The following is just meant to be an example - by no means it as introduction to dplyr. 

```{r}
require(dplyr)
aggr <- na.omit(summarize(group_by(big5,country),
                          E1_mean = mean(E1)))

# alternative syntax: 
big5 %>% 
  group_by(country) %>%
  summarize(E1_mean = mean(E1))


# use the revalue function from plyr.
# note that standard plyr can cause 
# trouble with dplyr sometimes, thus
# we only use ::
aggr$country <- plyr::revalue(aggr$country,
              c("US" = "USA",
                "DE" = "Germany",
                "CA" = "Canada",
                "CN" = "China",
                "ES" = "Spain",
                "FR" = "France")
              )


# subset to a dataset that only contains a few country with data
cs <- subset(aggr, country %in% c("USA","Germany","Canada",
                                  "China","Spain","France"))
cs$country <- factor(cs$country)

```

Now let's create a map of the world... 
Once we have done so, we need to merge our psychometric information to the map.


```{r}
# use the map_data function to have a standard 
# world map data.frame
world_map <- map_data("world")

data_map <- merge(world_map,cs,by.x = "region",
                  by.y = "country")

data_map_full <- merge(aggr,
                       world_map,
                       by.x = "country",
                       by.y = "region",all = T )



# built a plot layer by layer
g <- ggplot(data_map_full,aes(map_id = country)) +
  geom_map(map = world_map,color = "white") +
  geom_map(aes(fill = E1_mean), map = world_map,data = cs) +
  scale_fill_gradient(low="red", high="blue") +
  expand_limits(x = world_map$long, y = world_map$lat)

g


```

# Task3 
Execute the plot generated in object g line by line and explain what happens. 

# Factor Analysis in R


```{r}
items <- grep("^[A-Z]",names(big5),value = T)

test <- big5[,items]

test <- na.omit(test)

out <- factanal(test,5)

ei <- eigen(cor(test))
plot(ei$values,type = "b")

```






















